### 2020-10-26 thesis/__latexindent_temp.tex
Update index

% This template was initially provided by Dulip Withanage.
% Modifications for the database systems research group
% were made by Conny Junghans,  Jannik Strï¿½tgen and Michael Gertz

\documentclass[
     12pt,         % font size
     a4paper,      % paper format
     BCOR10mm,     % binding correction
     DIV14,        % stripe size for margin calculation
%     liststotoc,   % table listing in toc
%     bibtotoc,     % bibliography in toc
%     idxtotoc,     % index in toc
%     parskip       % p
aragraph skip instad of paragraph indent
     ]{scrreprt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% PACKAGES:

% Use German :
\usepackage[english]{babel}
% Input and font encoding
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
% Index-generation
\usepackage{makeidx}
% Einbinden von URLs:
\usepackage{url}
% Special \LaTex symbols (e.g. \BibTeX):
%\usepackage{doc}
% Include Graphic-files:
\usepackage{graphicx}
% Include doc++ generated tex-files:
%\usepackage{docxx}
% Include PDF links
%\usepackage[pdftex, bookmarks=true]{hyperref}

% Fuer anderthalbzeiligen Textsatz
\usepackage{setspace}

% hyperrefs in the documents
\usepackage[bookmarks=true,colorlinks,pdfpagelabels,pdfstartview = FitH,bookmarksopen = true,bookmarksnumbered = true,linkcolor = black,plainpages = false,hypertexnames = false,citecolor = black,urlcolor=black]{hyperref} 
%\usepackage{hyperref}

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

\usepackage{color,soul}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OTHER SETTINGS:

% Pagestyle:
\pagestyle{headings}

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}

\begin{document}

\listoftodos

% TITLE:
\pagenumbering{roman}
\begin{titlepage}
     \vspace*{1cm}
     \begin{center}
          \vspace*{3cm}
          \textbf
          {
               \Large University of Heidelberg\\
               \smallskip
               \Large Institute for Computer Science\\
               \smallskip
               \Large Working group database systems\\
               \smallskip
          }

          \vspace{3cm}

          \textbf{\large Bachelor thesis}

          \vspace{0.5\baselineskip}
          {
               \huge
               \textbf{Messaging Architecture for Integration of Customer Self-Services}
          }

     \end{center}

     \vfill
     {
          \large
          \begin{tabular}[l]{ll}
               Name:                 & Jonas Gann              \\
               Matriculation number: & 3367576                 \\
               Supervisor:           & Prof. Dr. Michael Gertz \\
               Date of submission:   & \today
          \end{tabular}
     }

\end{titlepage}

\onehalfspacing

\thispagestyle{empty}

\vspace*{100pt}
\noindent
I assure that I have written this bachelor thesis on my own and only used the specified sources and
resources and that I followed the principles and recommendations "Responsibility in Science" of the
University of Heidelberg.

\vspace*{50pt}
\noindent

\underline{\phantom{mmmmmmmmmmmmmmmmmmmm}}

\medskip
\noindent
Date of Submission: \today
\newpage

\chapter*{Zusammenfassung}

\newpage

\chapter*{Abstract}

\newpage

\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

\chapter{Context}
\todo[inline, caption={Expand Context Chapter}]{Add more content to this chapter}

Digital customer self-service (DCSS) is a great tool for improving customer experience and reducing cost. Many enterprises therefore heavily
rely on it for main business services. Especially recent governmental regulations like "Onlinezugangsgesetz" (OZG), "Datenschutzgrundverordnung"
(DSGVO) and "Digitale Versorgung Gesetz" (DVG) have increased requirements for DCSS capabilities.

DCSS is made available through DCSS providers. They can consist of websites and applications developed in self-work or be 
services provided by separate companies.

Adding DCSS to a system architecture is a common task. It usually involves integration of new DCSS components
into an existing system architecture. Due to the complicated nature of system architectures, integration proves to be a difficult
task.

\chapter{Objective}

\todo[inline, caption={Expand Objectives Chapter}]{Add more content to this chapter}

The objective of this bachelor thesis is to improve the capability of enterprises and organizations to integrate DCSS providers.
The thesis explains the basis of what DCSS and its providers are, which relevant scenarios exist and how DCSS providers can help
in the digital transformation.

The technological challenge of integrating DCSS providers into existing system architectures is focus of this bachelor thesis.

Based on scenarios resulting from governmental regulations (due to their current relevance), an integration architecture is presented, which describes
their integration. It provides information on which type of systems and data objects an existing
architecture requires. It models how CSS providers and system architectures can communicate -
deliver data and instructions. And it shows how, with fast development and deployment speed, integration can be applied to
heterogeneous system architectures, be non-invasive and reliable.

The integration architecture is evaluated in respect to technological feasibility and real-life applicability. The results are
incorporated into an operating manual, providing guidance in developing and deploying the presented integration architecture.

\chapter{Structure of Work}

\todo[inline, caption={Write Structure of Work Chapter}]{Briefly explain each chapter and their relationships}

\chapter{Basis}

\section{Digital Customer Self-Service (DCSS)}

\todo[inline, caption={Expand DCSS Chapter}]{Add more content to this chapter}

This chapter introduces the concept of digital customer self-service. It defines DCSS, describes what it is by usage of examples and mentions benefits such as challenges.

\textbf{Definition}: Digital self-service is the consumption of services provided by an enterprise over the internet without human to human interaction.

It is important to distinguish between "service" and "self-service". The difference is shown through two examples:

\textbf{Example 1}: A service is the availability to purchase items whereas digital self-service is the order placement over a website.

\textbf{Example 2}: A service is the availability of video on demand whereas digital self-service is the access of video files over a website.

Digital self-service is one of many possible ways to access a service: An order placement could also be done through email
or phone call and access to videos could be granted through same-day-delivery of DVDs.

Customers are able to do self-service because companies provide them with self-service tools. Creating them is not trivial
and is the main business model of many companies. Video on demand websites for example do not only flourish as a result of providing access to
videos (this business model existed long before them) but as a result of the digital self-service tools they provide.

Reasons for the widespread usage of digital self-service are significant benefits in ease of service usability for customers and cost reduction for enterprises.
Customers can self-service from almost every place, easily manage customer data like phone number and address, configure services like subscription plans and
access lots of information on available products or trouble shooting steps in an immediate way.
Digital customer self-service improves usability, saves time, increases availability, saves money and therefore increases customer experience
\cite[cf. 243]{CSSInternetAge}.
Enterprises, especially in progressive countries, aim to keep employment at a minimum, as it is a big cost factor. Digital self-service turns customers into
unpaid employees \cite{served}.

Decreasing employment cost while increasing customer experience sounds like a win-win situation. Sadly, digital self-service does have its downsides.
One downside is most immediate to the elderly. In order to use DCSS, computers have to be available and one has to be experienced in their usage.
Technological complexity, however, is not only a problem for customers but also for enterprises providing DCSS. Based on existing system architectures,
adding DCSS functionalities can turn out to be a difficult task. Some form of integration is often necessary. With increased system complexity and business
dependability come additional maintenance cost.

As tody most large enterprises provide DCSS, it can be assumed, that at a given size of an enterprise (measured in number of customers), the benefits of DCSS
surpass its disadvantages.

Governments see digitalization as a chance for improvements in many important social and economical areas\todo{references}. The process of digitalization can in
some cases be described as reinventing existing solutions for usability with computers. One of the goals is to make solutions more easily accessible for users and as a 
method, provide digital self-service. Governmental regulations relating to digitalization therefore often encourage or demand enterprises and organizations to 
provide DCSS.

\section{OZG}

\todo[inline, caption={Research Governmental Regulations Chapter}]{Explain the three regulations in general, what their purpose is what their challenges are}

\section{DCSS Provider}

\todo[inline, caption={Write CSS Providers Chapter}]{Explain what CSS Providers are (what services they can provide -> what IDAS does) and how they can help creating DCSS.
     Also explain the concept of a business connector as technological communication point of the CSS provider}

\subsection{Description}

\subsection{CSS Solutions}

\subsection{Business Connector}

\section{Enterprise Architecture}

\todo[inline, caption={Write Enterprise Architectures Chapter}]{Introduce Enterprise Architectures and challenges such as heterogeneous systems when trying to model them}

\begin{enumerate}
     \item Heterogeneous Enterprise Architecture Systems
           \begin{enumerate}
                \item Different Applications
                \item Different Application Vendors
                \item Different / No Application Interfaces
                \item Legacy Systems
           \end{enumerate}
     \item Different (proprietary) data models within and between Enterprises and Organizations
           \begin{enumerate}
                \item Different (property)name for same data objects (syntactic integration)
                \item Different meanings for same (property)name (semantic integration)
           \end{enumerate}
\end{enumerate}

\subsection{Modeling of the real world}

\subsection{Modeling Challenges}

\subsection{Enterprise Architecture Patterns (EAP)}

\section{Integration}

\todo[inline, caption={Write Integration Chapter}]{Introduce the concept of integration. What are challenges in general and what would be an ideal integration.
     Also explain the approach towards integration used (messaging, patterns)}

\begin{enumerate}
     \item Scarce Resources
           \begin{enumerate}
                \item Integration Development Speed
                      \begin{enumerate}
                           \item Necessary Development <=> Reuse of existing Technology
                           \item Complexity / Size of Integration
                      \end{enumerate}
                \item Maintenance of finished Integration
                \item Hardware / Software Costs of Integration
                      \begin{enumerate}
                           \item Licenses for Software
                           \item Scalability of Integration => Necessary Computing Power
                      \end{enumerate}
                \item Messaging Integration
                      \begin{enumerate}
                           \item Loose Coupling
                                 \begin{enumerate}
                                      \item Loose Coupling simplifies adaption to changing EA => simpler Maintenance
                                      \item Loose Coupling simplifies integration of new EA systems => integration of heterogeneous EA
                                      \item Loose Coupling allows Reuse of "Modules" => faster development
                                 \end{enumerate}
                           \item Messaging enables communication with many systems through Adapters
                           \item Stability of Integration
                                 \begin{enumerate}
                                      \item Future Changes of EA
                                      \item Scalability
                                      \item Failure of EA or Integration Components
                                 \end{enumerate}
                           \item Messaging provides mechanisms for Stability
                                 \begin{enumerate}
                                      \item Store-and-Forward
                                      \item Load Balancing
                                 \end{enumerate}
                      \end{enumerate}
                \item Integration Patterns
                      \begin{enumerate}
                           \item Patterns speed up construction of Integration Architecture
                           \item Patterns are proven solutions
                           \item Patterns abstract from concrete technologies
                                 \begin{enumerate}
                                      \item Simplifies understanding of integration concept
                                      \item Allows implementation with different technologies
                                 \end{enumerate}
                      \end{enumerate}
           \end{enumerate}
\end{enumerate}

\subsection{Definition}

\subsection{Requirements}

\subsubsection{Loose Coupling}

\subsubsection{Homogeneous Landscapes}

\chapter{Preparatory Work}

\section{OZG Scenarios}

\todo[inline, caption={Research DCSS Scenarios Chapter}]{
     \begin{enumerate}
          \item Should governmental regulations be the focus of scenarios? => good resources for OZG
     \end{enumerate}
}

Usage of DCSS OZG scenarios (\textbf{process diagrams and data schemas available}
\cite{ozgScenarios}), DSGVO scenarios (textual description available \cite{dsgvo}) and DVG scenarios
(textual desciption available \cite{gematik-e-rezept})
\\\\
The scenarios include more than just CSS and are partly non-digital. Therefore processes relevant for DCSS have to be selected
and it has to be defined which processes are part of a CSS provider, the business systems or the integration architecture:

\begin{enumerate}
     \item What part of a scenario is digital?
     \item What part of a scenario can be done through a DCSS provider / is relevant for DCSS?
     \item What part of a scenario is done through \textbf{existing} business systems?
           \begin{enumerate}
                \item new necessary business logic part of (configurable) integration? => non invasive
           \end{enumerate}
     \item What part of a scenario should be done by the integration architecture?
     \item What are commonalities of scenarios?
           \begin{enumerate}
                \item OZG scenarios often describe digital handling of applications
                \item In OZG scenarios, often multiple governmental institutions are involved. This could be modeled as one system architecture with the problem of distributed
                      data and functionality which the integration architecture would solve.
           \end{enumerate}
\end{enumerate}

\begin{itemize}
     \item Unemployment Benefit (AG2)
\begin{itemize}
     \item Data Schema
     \item Process Diagram
     \item Process Diagram of Application
     \item Click-Dummy
\end{itemize}

\item Training Promotion (BAföG)
\begin{itemize}
     \item Data Schema
     \item Process Diagram
     \item Click-Dummy
\end{itemize}

\item Drivers License
\begin{itemize}
     \item Data Schema
     \item Process Diagram
     \item Process Diagram of Application
     \item Click-Dummy
\end{itemize}

\item Company Registration and Approval
\begin{itemize}
     \item Data Schema
     \item Process Diagram
     \item Click-Dummy
\end{itemize}

\item Visa Issuance
\begin{itemize}
     \item Data Schema
     \item Process Diagram
\end{itemize}

\end{itemize}

\section{Relevant Systems and Data}

\todo[inline, caption={Research Relevant Systems and Data Chapter}]{Explain the approach of modeling enterprise architectures by EAPs and explain the used EAPs.
     Analyze which system and data bricks of the system architecture could be necessarry for DCSS}

\subsection{EAPs}

\subsection{Architecture Bricks}

\subsection{Data Bricks}

\subsection{Integration Requirements}

\subsubsection{Regarding Integration Architecture}

\subsubsection{Regarding Business Connector}

\section{Business Connector}

\todo[inline, caption={Research Business Connector Chapter}]{Construct the model of the business connector}

\subsection{Functionalities and Interfaces}

\subsection{Integration Requirements}

\subsubsection{Regarding Enterprise Architecture}

\subsubsection{Regarding Integration Architecture}

\subsection{Documentation}

\subsubsection{Connector as Architecture Brick}

\chapter{Integration Architecture}

\todo[inline, caption={Integration Architecture Chapter}]{Create the integration architecture}

\section{Scenario 1}

\subsection{Integration Documentation}

\subsection{Required System and Data Bricks}

\subsection{System Integration}

\subsection{Data Integration}

\section{Scenario 2}

\subsection{Integration Documentation}

\subsection{Required System and Data Bricks}

\subsection{System Integration}

\subsection{Data Integration}

\chapter{Integration Architecture Evaluation}

\section{Technology}

\section{Customer Example}

\section{Operating Manual}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}