### 2021-01-28 sections/3-structure_of_work.tex

The "Fundamentals" chapter explains terminology and basic concepts.

In the following chapter called "Identity Management System" the thesis analyzes requirements for identity management through the perspective of the user and through the perspective of service providers for the OZG. It is presented, how an identity management provider could solve current identity management problems and which challenges for an integration exist.

The chapter "Integration Proposal" first documents a possible IMP system and connector, based on the requirement analysis of the previous chapter. It then presents the integration architecture with iteratively increasing requirements through textual description, multiple flow diagrams and messaging architectures.

A "Solution Evaluation" analyzes benefits and problems of the proposed solution.

The "Outlook" will conclude the thesis by presenting possibilities for advanced integration architectures.