### 2021-01-28 sections/6-requirement_analysis.tex

\subsection{User Perspective}

\begin{itemize}
    \item The user wants to manage his personal information in one place
    \begin{itemize}
        \item The user wants to CRUD attributes
        \item The user wants SPs to CRUD his attributes
        \item The user wants a user friendly interface where he can CRUD his attributes
        \item In case the user cannot understand an attribute or its possible values, he needs OSS tools which enables him to enter the correct value
        \item The user wants to enter personal information only once (=> SPs should reuse existing attributes)
        \item The user wants a new or modified attributes to be instantly used everywhere
        \item The user wants a deleted attribute to be instantly deleted everywhere
        \item The user wants to have a list of often used attributes, in order to prefill them
    \end{itemize}
    \item The user wants to manage authentication in one place
    \begin{itemize}
        \item The user does not want to authenticate
        \item The user wants that only he has access to the identity
    \end{itemize}
    \item The user wants to manage authorisation in one place
    \begin{itemize}
        \item The user wants to manage who can perform which actions in his name (f.e CRUD attributes, place order, cancel subscription)
        \item The user wants an authorisation to only be granted if he manually approves
        \item The user wants an authorisation to be associated with a single entity, a reason, which actions can be performed and duration of the authorisation
        \item The user wants to communicate with the authorised entity
        \item The user wants to have much information about the authorised entity
        \item The user wants to be able to revoke an authorisation
        \item The user wants to save the history of authorisations
    \end{itemize}
    \item The user wants to use all OSS tools in one place
    \begin{itemize}
        \item Communication through inbox
        \item Subscribe for a service of an SP
        \item Place order in a shop of an SP
        \item ...
    \end{itemize}
    \item The user wants

\end{itemize}

\subsection{Service Provider Perspective}

\begin{itemize}
    \item SPs want to manage personal information of users
    \begin{itemize}
        \item SPs want to CRUD attributes of the user
        \item In case modification of an attribute has strong implications on the operation of a system, the modification of an attribute needs to be limited
        \item SPs want to access personal information already collected by other SPs
        \item SPs want 
    \end{itemize}
    \item identification
    \item authentication
    \item authorisation
    \item 
\end{itemize}

\subsection{OZG Perspective}