### 2021-02-06 sections/6-requirement_analysis.tex

\subsection{User Perspective}

\begin{itemize}
    \item The user wants one service where he can manage all personal information

    \item The user does not want to manage multiple user profiles

    \item The user wants all service providers to be able to use the personal information of the service

    \item The user wants service providers to always use the most recent state of his personal information

    \item The user wants to be able to share and retract personal information, service providers have access to

    \item The user wants to restrict access to personal information as much as possible

    \item The user wants information about the service provider he is about to share information with

    \item The user wants to know the reason each time a service provider wants to process his personal information

\end{itemize}

\subsection{Service Provider Perspective}

The service provider is assumed to have an existing system architecture containing systems for management of user profiles. This section describes requirements of a service provider regarding the service for managing user data.

\begin{itemize}

    \item The service provider wants a good user experience -> see all user requirements above

    \item The service provider wants to access personal information necessary for processing user requests

    \item The service provider wants to change, add or delete personal information

    \item The service provider wants to communicate with users

    \item The service provider wants to the service to guarantee secure interaction with users

    \item The service provider wants the system to keep a record of interactions with users and be able to verify the validity of the record

    \item The service provider wants to inform users about legal regulations like privacy protection measurements

    \item The service provider wants minimal integration effort and cost

    \item The service provider wants to be able to quickly integrate parts of the service with optional future upgrades

    \item The service provider wants the integrated solution to be scalable, maintainable and stable

\end{itemize}