### 2021-02-21 sections/12-advanced_utilization.tex

The goal of this integration architecture is to make as much functionalities required for the basic OZG use case available through the IMP client. Now, the OZG user profiles are replaced with IMP identities and the integration architecture integrates with any system component necessary.

\section{IMP Solution Integration}

\subfile{13-imp_solution_integration}

\section{Technological Integration}

\subfile{14-technological_integration}