### 2021-02-21 sections/14-technological_integration.tex

\begin{center}
    \includegraphics[scale=0.6]{Diagrams/Integration Architecture 2/Integration Overview.pdf}
\end{center}