### 2021-03-04 sections/12-advanced_utilization.tex

The goal of the advanced IMP utilization for OZG is to improve upon the solution presented in chapter 4 while allowing it to be modify and replace large portions of the existing system architecture.

\section{IMP Solution Integration}

\subfile{13-imp_solution_integration}

\section{Technological Integration}

\subfile{14-technological_integration}