### 2021-03-06 chapters/2-background_and_related_work.tex
überarbeitung

This chapter covers three important topics, following chapters rely on.

Identity Management Provisioning is introduced as a new identity management method. Its features along with system components and important use cases are described.

The Online Access Law is introduced as a currently relevant regulation with special requirements for identity management. The most important regulations as well as resulting system architectures for its implementation are described.

Messaging is introduced as a way of designing integration architectures. Multiple messaging patterns as well as their visual notation are described.

\section{Identity Management Provisioning (IMP)}
\subfile{sections/5-imp}

\section{Online Access Law (OZG)}
\subfile{sections/6-online_access_law}

\section{Messaging}
\subfile{sections/7-messaging}