### 2021-03-06 chapters/sections/15-conclusion.tex

Finally, this chapter aims to summarize the most important results of the thesis and give an outlook on possible future developments and contributions.

\section{Summary}

\begin{itemize}
    \item Purpose of the thesis is to design integration solutions for identity management provisioning systems on the example of OZG.
    \item The integration is separated into a business scenario integration trough an IMP solution and a system integration trough a messaging system.
    \item Trough two IMP solutions, it was shown, how IMP can be utilized in different identity management scenarios
    \item In the first scenario, an IMP solution was presented which focused on enhancing usability of 
    \item The thesis introduced identity management provisioning as an identity management service capable of replacing conventional user profiles
    \item On the example of the OZG, the thesis showed 
\end{itemize}

\section{Outlook}

\begin{itemize}
    \item experimental evaluation
    \item 
\end{itemize}