### 2021-03-13 chapters/sections/1-context.tex
Anfang und ende fertig

Many services, are accessible through the internet. Especially during the COVID-19 pandemic, stores had to switch to online trade due to lockdowns. Online Service Providers (SP) often require customers to create user profiles in order to access their services. Management of personal information as well as interaction and communication between Service Provider and customer are often only possible through a web page or e-mail. It is time consuming to maintain even a small amount of user profiles: To change personal information, like an e-mail address, the web page of each Service Provider has to be accessed the correct credentials have to be submitted and the information has to be manually filled in each time.

In some cases, user profiles are required by Service Providers even if their service could be provided without them. Especially in case of small stores, running their individual online shop, user profiles are not necessary for customers to place an order. Personal information can be shared temporarily until processing of the order is completed and the package delivered. Some online shops allow users to place orders by temporarily sharing personal information, which however, limits interaction and communication possibilities to insecure e-mail exchange.

A method for identity management called Identity Management Provisioning (IMP) is introduced. IMP provides identity management as a service to users and Service Providers. It enables users to create and manage a single digital identity and use it to share personal information and interact with multiple Service Providers through "IMP relationships" and "IMP messages". Through IMP relationships, users and Service Providers have a tool to securely share personal information which, depending on the use case can be used in different ways. It can be used to create and manage a persistent user profile or to temporarily share personal information for a business process. IMP messages enable user and Service Provider to securely exchange structured data and depending on the client application which receives and displays the data, design rich user interactions.