### 2021-03-13 chapters/sections/3-structure_of_work.tex

In the chapter "Background and Related Work" important topics that following chapters rely on are covered.
In the chapter "User profile Identity Management" disadvantages of conventional user profile based identity management are described as well as possible solutions through IMP.
In the chapter "Enhancing Utilization in OZG" the first IMP solution is presented as well as its integration through the IMPS integration architecture.
In the chapter "Replacing IMP Utilization" the second IMP solution is presented as well as its integration through the IMPS integration architecture.
In the chapter "Conclusion" the most important results of the thesis are summarized and possible future developments and contributions are discussed.