### 2021-03-13 chapters/sections/1-context.tex

Today, many services are accessible through the internet. Especially during the COVID-19 pandemic, stores had to switch to online trade due to lockdowns. Online Service Providers (SP) often require customers to create user profiles in order to access their services. Management of personal information as well as interaction and communication between Service Provider and customer are often only possible through a web page or e-mail. It is time consuming to maintain even a small amount of user profiles: To change personal information, like an e-mail address, the web page of each Service Provider has to be accessed the correct credentials have to be submitted and the information has to be manually filled in each time.

In some cases, user profiles are required by Service Providers even if they are not necessary. Especially in case of small stores, running their individual online shops, user profiles are not necessary for customers to place an order. Personal information can be shared temporarily until processing of the order is completed and the package delivered. Some online shops allow users to place orders by temporarily sharing personal information, which however, limits interaction and communication possibilities to insecure e-mail exchange.

A currently relevant topic with an interesting approach for identity management is the German Online Access Law (Online Zugangs Gesetz - OZG). It requires German federal republic, each federal state and commune to provide administration portals for digital access to administrative services. Federal republic and each member state operate their individual user profile systems accessible through their administration portals.

The Online Access Law therefore constructs a system landscape which demonstrates the scenario of distributed partial identities on a small scale while maintaining the same problems as the private sector: Users have to create, maintain and login into multiple user profiles to access administrative services distributed across administration portals. To solve this problem, a solution for making OZG user profiles "interoperable" is in development. However, interoperable user profiles can not be accessed by the private sector.

Identity Management Provisioning (IMP) is an approach to identity management which aims to solve problems of distributed partial identities by enabling users to bring their online identities with them, when accessing systems of Service providers. It provides identity management as a service to users and Service Providers by enabling users to create and manage a single digital identity and use it to share personal information and interact with multiple Service Providers through "IMP relationships" and "IMP messages". Through IMP relationships, users and Service Providers have a tool to securely share personal information which, depending on the use case can be utilized in different ways. It can for example be used to create and manage a persistent SP user profile or to temporarily share personal information for a business process of a SP. IMP messages enable user and Service Provider to securely exchange structured data and depending on the client application which receives and displays the data, design rich user interactions. This approach to identity management could be an alternative to user profiles in the private sector and in context of the OZG.

Identity management systems do not stand alone but have to be accessible by various other systems in a system architecture to for example access personal information and perform authentication. They also play an important role in the design of business processes. Therefore, for Service Providers to utilize IMP, integration into existing business processes and existing system architectures is necessary.