### 2021-03-13 chapters/sections/3-structure_of_work.tex

The chapter "Background and Related Work" covers important topics that following chapters rely on. 
The chapter "User Profile Identity Management" describes disadvantages of user profile based identity management in general as well as in context of the OZG. Possible solutions through IMP are described. 
The chapter "IMP Utilization as Extension" presents the first IMP solution as well as its integration through the technological integration architecture.
The chapter "IMP Utilization as Replacement" presents the second IMP solution as well as its integration through the technological integration architecture.
The chapter "Conclusion" summarizes the most important results of the thesis and mentions possible future developments and contributions.