### 2021-03-26 chapters/2-background_and_related_work.tex
finished

This chapter covers three important topics that the following chapters rely on. Identity Management Provisioning as an approach to solve problems of distributed partial identities is described along with possible services and systems. The most important regulations of the Online Access Law as well as its basic use case and the resulting system architecture of a federal state are described. Messaging is presented as a way of designing technological integration architectures. Multiple messaging patterns as well as their visual notation are explained.

\section{Identity Management Provisioning (IMP)} \label{imp}
\subfile{sections/5-imp}

\section{Online Access Law (OZG)} \label{ozg}
\subfile{sections/6-online_access_law}

\section{Messaging} \label{messaging}
\subfile{sections/7-messaging}